AA: the Algoritmic Autoregulation Software Development Methodology

### 2013-05-31 paper.tex
feita correcao da intro

One of the defining features of modern times is the widening
geographical distribution of software teams~\cite{last2003}.
This is responsible for the so called Global Software Development (GSD)~\cite{german2003}. 
An example is the free software
movement. Projects and institutions like Mozilla Foundation has several
employees and thousands of voluntary developers distributed across
many countries. The same is true for GNOME~\cite{german2003}, OpenBSD,
MySQL or Apache Software Foundation, to cite just a few of the most active
projects.\footnote{Ohloh, the open source network, have a more complete
  and constantly updated list of the most active projects on-line at
  \url{www.ohloh.net}} Along the free and open software
projects, GSD is growing popularity in every niche of the software
industry as a whole, even on those distributing their software with
proprietary licenses. This phenomenon is attributed to a variety of factors
such as a larger labor pool, natural globalization of software companies
and foundations or even the premise of cheaper cost of
production~\cite{komi2005}.

However, dispite the advantages of GCD, we have noticed how
difficult it is to coordinate and fund free software on a larger scale
than currently available, when teams are very heterogeneous containing
not only volunteers and very experienced developers, but also
contractors from different backgrounds and cultures. Our observations
found base on the factors suggested by Carmel~\cite{carmel1999} as
main difficulties for GCD: distance, time and cultural differences. In
the case of free or open software projects, all these factors are
involved.

Another problem faced by modern software companies and other
collectives are frequent ineffective meetings, which are seldom
focused to the interest to any attendant. The result is that it has
become the norm to participate in too many meetings with the ``laptop
open'', which is unproductive at the very least. Software developers
like to code, to be productive, to have their hands on their project,
to do what they are best at. They dislike to have to stop for meetings
or to write lengthy reports to justify their funding.
%\todo{ler mais. cacm, etc}

Here we propose AA, a methodology and an associated software system
for coordinating distributed team work dealing with the disadvantages
of GSD. Team members take on an egalitarian role, and stay
voluntarily \textit{logged} in the system for part of their time
(e.g.\ 2 hours per day), during which they log a periodical short text
sentence --- similar to a ``tweet'' in this Twitter era --- as the
status of their activity using an easy to use command. These
``microblog sentences'' are publicly aggregated and validated by other
team members. Through AA, we have a methodology and an associated
system to validate and enable the activities of a distributed software
team. It implicitly legitimizes financial support for the expansion of
the activity of the developer team. The AA methodology is specially
useful for coordinating distributed and decentralized team work,
providing effective means to asynchronously update different team
members without the need for synchronous unproductive meetings.

@@

A brief overview of current papers about GSD methodologies related with
AA is presented in Section~\ref{related-work}, while in
Section~\ref{aa-methodology} we describe the most relevant
characteristics of the AA methodology followed by the description of
our experience using AA in a team of 9 developers since July of
2011. Finally, in Section~\ref{conclusions}, we draw some final
conclusions and indicate future possibilities to the
practical use of AA on other teams of software developers or
individuals working on non-software activities.

@@

%% brief introduction
In a scenario where GCD is growing as a popular form of software
development, not just on free software projects but in the whole
software industry, we need methodologies to deal with its
disadvantages and at the same time to amplify its advantages.

This paper has presented a methodology to GCD, being the development
conducted on large or small groups of software developers, working on
different countries or even at the same room. The AA methodology
implements a simple system where each developer take notes of his work
generating a periodical log of small text sentences. The sum of those
sentences, along an entire session of work, results in a complete
report. The report is made public available through a Website and be
validated by other developers sorted aleatory by the AA Web server.

>>>>

One of the defining features of modern times is the widening
geographical distribution of software teams~\cite{last2003}.
This is responsible for the so called Global Software Development (GSD)~\cite{german2003}. 
An example is the free software
movement. Projects and institutions like Mozilla Foundation has several
employees and thousands of voluntary developers distributed across
many countries. The same is true for GNOME~\cite{german2003}, OpenBSD,
MySQL or Apache Software Foundation, to cite just a few\footnote{Ohloh, the open source network, have a more complete
  and constantly updated list of the most active projects on-line at
  \url{www.ohloh.net}}. Along the free and open software
projects, GSD has a growing popularity in every niche of the software
industry as a whole, even on those distributing their software with
proprietary licenses. This phenomenon is attributed to a variety of factors
such as a larger labor pool, natural globalization of software companies
and foundations or even the premise of cheaper cost of
production~\cite{komi2005}.

Dispite the advantages of GSD, it is known how
difficult it is to coordinate and fund free software on a larger scale
than currently available. That is, when teams are very heterogeneous, containing
not only volunteers and experienced developers, but also
contractors from different backgrounds and cultures. This observations
are in consonance with dedicated studies, such as~\cite{carmel1999}.
As a common base, related difficulties for GSD gather around distance, time and cultural differences. In
the case of free or open software projects, all these factors are
involved.

Another problem faced by modern software companies and other
collectives are frequent ineffective meetings, which are seldom
focused on the interest of any attendant. The result is that it has
become the norm to participate in these with the ``laptop
open'', which is unproductive. Software developers
like to code, to be productive, to have their hands on their project,
to do what they are best at. They dislike to have to stop for meetings
or to write lengthy reports to justify their funding.
%\todo{ler mais. cacm, etc}

To address these matters is the purpose of AA, a methodology and an associated software system
for coordinating distributed team work dealing with the disadvantages
of GSD. Team members take on an egalitarian role, and stay
voluntarily \textit{logged} in the system for part of their time
(e.g.\ 2 hours per day), during which they log a periodical short text
sentence --- similar to a ``tweet'' in this Twitter era --- as the
status of their activity using an easy to use command. These
``microblog sentences'' are publicly aggregated and validated by other
team members. Through AA, the community has a methodology and an associated
system to validate and enable the activities of a distributed software
team. It implicitly legitimizes financial support for the expansion of
the activity of the developer team as the activities are documented and validated.
The AA methodology is specially
useful for coordinating distributed and decentralized team work,
providing effective means to asynchronously update different team
members, diminishing the need for harsh synchronous meetings.

@@

A brief overview of current papers about GSD methodologies related with
AA is presented in section~\ref{related-work}. In
section~\ref{aa-methodology}, the most relevant
characteristics of the AA methodology are outlined. In this same section,
there is a description of
an AA use in a team of 9 developers in the second half of
2011 and a broader a more released use of AA from 2012 until now. Finally, in Section~\ref{conclusions}, there are final
conclusions and indicatives of future possibilities to the
practical use of AA on other teams of software developers or
individuals working on non-software activities.

@@

%% brief introduction
In a scenario where GSD is growing as a popular form of software
development, not just on free software projects but in the whole
software industry, we need methodologies to deal with its
disadvantages and at the same time to amplify its advantages.

This paper has presented a methodology to GSD, being the development
conducted on large or small groups of software developers, working on
different countries or even at the same room. The AA methodology
implements a simple system where each developer take notes of his work
generating a periodical log of small text sentences. The sum of those
sentences, along an entire session of work, results in a complete
report. The report is made public available through a Website and be
validated by other developers sorted aleatory by the AA Web server.
